\documentclass[11pt]{NSFamsart}
\usepackage{latexsym,amsfonts,amsmath,epsfig,multirow}
\usepackage{stackrel,tabularx,mathtools,enumitem,xspace}
\usepackage[dvipsnames]{xcolor}
\usepackage[numbers]{natbib}
\usepackage{hyperref,accents}
\usepackage[margin=1in]{geometry}

% This package prints the labels in the margin
%\usepackage[notref,notcite]{showkeys}


%\pagestyle{empty}
\thispagestyle{plain}
\pagestyle{plain}

%\headsep-0.6in
%\headsep-0.45in

\begin{document}
%\setlength{\leftmargini}{2.5ex}

\centerline{\textbf{\Large Budget Justification}}

\subsection*{Salaries and Wages}

The PIs are budgeted at a modest one month of summer salary each per year.  Fred J. Hickernell, Principal Investigator and Gregory E. Fasshauer, Co-Principal Investigator jointly lead the research.  FJH will be involved more on Research Projects 2 and 3, and the Co-PI will be involved more on Research Projects 1, 3, and 4.  Together, they lead the weekly research seminar and they meet individually with the students and visitors working on the research projects most closely related to them.

The proposed salaries for graduate research assistants support only a small number of the PhD students working under our supervision on projects related to this grant.  Others are teaching assistants or are self-supported.  We direct RA support to those students who have the potential to make the most contributions.

The undergraduate salaries are for two REU students per summer.

\subsection*{Fringe Benefits}
Fringe benefit rate is 22.3\% for academic year salary and 7.70\% for the summer month
salary.

\subsection*{Travel}
The PIs travel to several conferences each year to disseminate their work and to learn what others have discovered.  These include, e.g., the Joint Mathematics Meetings, SIAM annual meetings, Midwest Numerical Analysis Days, the biennial Monte Carlo and Quasi-Monte Carlo conferences, as well as the triennial Dolomite Workshops on Constructive Approximation and Applications, International Conferences on Approximation Theory, and Workshops on Multivariate Approximation and Interpolation with Applications. Although our travel is sometimes sponsored by the organizers, that is not always the case.

We often take our students to conferences with us so that they may have the experience of presenting their work and networking with the academic community.  The REU students are always encouraged to attend conferences to present their work.  Although the organizers sometimes have funding for young scholars, this does not cover all costs.

Besides the substantial travel by the PIs and our students, we have asked for significant additional funds to support our two senior persons, Drs. Sou-Cheng Choi and Michael McCourt.  They will be heavily involved in this project, as they have been in our previously funded work.  Each year we want to sponsor Dr. Choi's travel to conferences, and we want to support Dr. McCourt's travel to Chicago for collaboration.

\subsection*{Supplies}
Costs for licenses for specialized mathematical software for research purposes.

\subsection*{Other Direct Costs - Tuition}
The current tuition rate is \$1,250 per credit hour. Amount requested is 2 students $\times$
\$1,250 $\times$ 9 credit hours per year = \$25,000 per year.

\subsection*{Indirect Costs}
Our federally negotiated rate is 52\% of modified total direct costs.

A 4\% annual increase applies to all categories

\end{document} 