% !TEX TS-program = PDFLatexBibtex
%&LaTeX
%Greg and Fred's NSF Grant Dec 2014
\documentclass[11pt]{NSFamsart}
\usepackage{latexsym,amsfonts,amsmath,epsfig,multirow,stackrel,natbib,tabularx,enumitem,xspace}

% This package prints the labels in the margin
%\usepackage[notref,notcite]{showkeys}


%\pagestyle{empty}
\thispagestyle{plain}
\pagestyle{plain}

\headsep-0.6in
%\headsep-0.45in

\setlength{\textwidth}{6.5in}
\setlength{\oddsidemargin}{0in}
\setlength{\evensidemargin}{0in}
\textheight9in
%\textheight9.1in



%\setcounter{page}{1}


\begin{document}
%\setlength{\leftmargini}{2.5ex}

\centerline{\large \bf Facilities, Equipment and Other Resources}

\bigskip

Sou-Cheng Choi and Michael McCourt are two collaborators, unfunded by this proposal except for travel support. 

All IIT faculty, PhD students, and visitors have offices provided at IIT.  Summer MS, BS, and high school students have shared work areas.  The Department of Applied Mathematics has a research computer room that is available to all members of our department.  

The Department of Applied Mathematics has a 128-core cluster named Karlin funded by NSF-DMS-0923111 Scientific Computing Research Environments for the Mathematical Sciences at IIT, which is available for this project to use.  Karlin is housed in the IIT Research Data Center, a secured, monitored data storage service available to all IIT research faculty, centrally managed by IIT Office of Technology (OTS) Services.

IIT has site licenses for Mathematica, MATLAB, SAS, and JMP.

IIT’s university library provides excellent access to journals, research monographs, and databases, either on-site, online or via inter-library loan.

IIT was listed on the National Federal Register of Historic Places in 2005. The proposed research activities will not make any physical changes to IIT’s campus and buildings.





\end{document}
